Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-24 paper.tex
sec 2

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machine involves two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an
$\sOOinspect{e}{\rho}{\kappa}$ or $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration.

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a function in cast, the machine first casts $v_1$, the
operand, then applies the cast operand to $v_2$, the underlying
function, and finally cast the return value.
%
To take out the first (resp. second) part of a pair in cast, the
machine firstly take out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
%In case splitting, if the target value is a left injection, the
%machine moves to a state that will apply the first continuation
%function to the value inside the left injection. The case for right
%injection is similar.
%
To case split an injection in cast, the machine moves the cast from
the injection to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value.  If the
cast succeeds, the machine returns the result to the next
continuation.  If the cast fails, the machine halts with the blame
label.
%
%% When the continuation is $stop$, the machine enters a halting
%% state. The When the machine is in a halting state, it stays in the
%% same state.

The reflexive transitive closure of reduction ($\judgeCreduceTrans{s}{s}$) and 
evaluation ($eval_\mathcal{x}(e)$)
are standard~\citep{felleisen03:_pllc}.

@@

Let $v$ range over values. A value is either the unit, a function, a pair, a 
left injection, a right injection, or a value in cast. 
As shown by the value typing rules, for values in casts, 
if the cast is between pre-types, source and target types must have the 
same type constructor at the top. 
If the cast is from a pre-type to the dynamic type, the pre-type must be 
injectable. The definition of injectable types depends on blame
strategies: for the \lazyD\ strategy, every pre-type is injectable;
for the \lazyUD\ strategy, a pre-types is injectable if all its sub-parts are 
the dynamic type.
%As shown by the value typing rules, if the target type of a
%value in cast is the dynamic type, the underlying value must be of an
%injectable type. The definition of injectable types depends on blame
%strategies: for the \lazyD\ strategy, every pre-type is injectable;
%for the \lazyUD\ strategy, a pre-types is injectable if and only if
%all its sub-parts are the dynamic type. If the target type of a value in cast 
%is a pre-type, the type of the underlying value must have the
%same type constructor.

Let $c$ range over casts. A cast is a triple of a type, a label, and a type.

Let $o$ range over observations. They are what we would observe if 
a program terminates. Observations include all value constructors and
blame.  The function converting values to observations ($observe(v) =
o$) is defined in the natural way.

@@

Let $r$ range over cast results. A cast result is either a success, which 
brings back a value, or a failure, which brings back a blame label.

Let $s$ range over machine states. A state is either looking at an 
expression to decide what to do next, returning a value to a continuation, or 
halting with an observation.

Let $\kappa$ range over continuations. $\mathtt{stop}$ is the top 
continuation. The remaining continuations correspond to expressions. For 
example, $\kOOconsI{T_1}{T_2}{e}{\rho}{\kappa}$ is the continuation waiting for 
the first part of the pair. And the last continuation, $\kOOcast{c}{\kappa}$ is 
to cast the value before returning to $\kappa$.

@@

Let $v$ range over values. In \ineffCEK{} we have one value
constructor for all values in casts.  In \effCEK{C}, however, we do not
have this generic value constructor, instead we push those casts into
the ordinary values. For instance, $(\vOOcons{v_1}{v_2})$ corresponds
to $(\hcvOOcons{v_1'}{id(T_1)}{v_2'}{id(T_2)})$. To construct values
of the dynamic type, we use $\hcvOOinj{P}{v}$.

@@

\todoKC{write conclusion. -KC}

@@

>>>>

Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
machine and Fig.~\ref{fig:state} gives a grammar for machine states
$s$, including a definition of values and value typing. 
%Fig.~\ref{fig:state} gives a grammar for machine states
%$s$, including a definition of values and value typing and
%Fig.~\ref{fig:machine-cekc} defines the transition relation of the CEK
%machine.
The transitions involving casts are highlighted in red and described in
more detail below. The other transitions are standard for a CEK
machine for an extended simply typed lambda calculus.
%
Recall that CEK machine involves two kinds of transitions, (1) those
that dive further into an expression (looking for a redex) and push an
entry onto the continuation, and (2) those that return a value to the
current continuation and possibly perform a computation.
Corresponding to (1) and (2), the machine state is either in an
$\sOOinspect{e}{\rho}{\kappa}$ or $\sOOreturn{v}{\kappa}$ configuration, 
respectively. Additionally, there is the $\sOOhalt{o}$ configuration, where the 
machine halts with an observation.
%Let $s$ range over machine states. A state is either looking at an 
%expression to decide what to do next, returning a value to a continuation, or 
%halting with an observation.

Let $v$ range over values. A value is either the unit, a function, a pair, a 
left injection, a right injection, or a casted value. Value typing rules 
restrict the casts in casted values. If the cast is between pre-types, source 
and target types must have the same type constructor at the top ($P_1 \approx 
P_2$). If the cast is from a pre-type to the dynamic type, the pre-type must be 
injectable. The definition of injectable types depends on blame
strategies: for the \lazyD\ strategy, every pre-type is injectable;
for the \lazyUD\ strategy, a pre-types is injectable if all its sub-parts are 
the dynamic type.

The transition relation $\judgeCreduce{s}{s}$ is parameterized over
$applyCast$ to allow for the differences between D and UD.
%
When evaluating a cast expression, the machine moves the cast to the
continuation and evaluates the inner expression.
%% Other
%% transition rules starting from$Eval$states are standard.
To apply a function in cast, the machine first casts $v_1$, the
operand, then applies the cast operand to $v_2$, the underlying
function, and finally cast the return value.
%
To take out the first (resp. second) part of a pair in cast, the
machine firstly take out the first (resp. second) part of $v$, the
underlying pair, and cast the result.
%
%In case splitting, if the target value is a left injection, the
%machine moves to a state that will apply the first continuation
%function to the value inside the left injection. The case for right
%injection is similar.
%
To case split an injection in cast, the machine moves the cast from
the injection to continuations functions.
%
To cast a value, the machine invokes $applyCast$ on the value. If the cast 
succeeds ($\rOOsucc{v'}$), the machine returns the result to the next
continuation.  If the cast fails ($\rOOfail{l}$), the machine halts with the 
blame label.
%Let $r$ range over cast results. A cast result is either a success, which 
%brings back a value, or a failure, which brings back a blame label.

%
%% When the continuation is $stop$, the machine enters a halting
%% state. The When the machine is in a halting state, it stays in the
%% same state.

The reflexive transitive closure of reduction ($\judgeCreduceTrans{s}{s}$) and 
evaluation ($eval_\mathcal{x}(e)$) are standard~\citep{felleisen03:_pllc}.
Let $o$ range over observations. They are what we would observe if 
a program terminates. Observations include all value constructors and
blame.  The function converting values to observations ($observe(v) =
o$) is defined in the natural way.

@@

%As shown by the value typing rules, if the target type of a
%value in cast is the dynamic type, the underlying value must be of an
%injectable type. The definition of injectable types depends on blame
%strategies: for the \lazyD\ strategy, every pre-type is injectable;
%for the \lazyUD\ strategy, a pre-types is injectable if and only if
%all its sub-parts are the dynamic type. If the target type of a value in cast 
%is a pre-type, the type of the underlying value must have the
%same type constructor.

@@

%Let $\kappa$ range over continuations. $\mathtt{stop}$ is the top 
%continuation. The remaining continuations correspond to expressions. For 
%example, $\kOOconsI{T_1}{T_2}{e}{\rho}{\kappa}$ is the continuation waiting 
%for 
%the first part of the pair. And the last continuation, $\kOOcast{c}{\kappa}$ 
%is 
%to cast the value before returning to $\kappa$.

@@

Let $v$ range over values. In \ineffCEK{} we have one value
constructor for all castes values.  In \effCEK{C}, however, we do not
have this generic value constructor, instead we push those casts into
the ordinary values. For instance, $(\vOOcons{v_1}{v_2})$ corresponds
to $(\hcvOOcons{v_1'}{id(T_1)}{v_2'}{id(T_2)})$. To construct values
of the dynamic type, we use $\hcvOOinj{P}{v}$.

@@

@@

In this
